---
title: "UNL01"
format: pdf
editor: visual
---

```{r setup, include=FALSE}
library(tidyverse)
library(afex)
library(BayesFactor)
library(apa)
library(emmeans)
library(rstatix)
options(scipen = 999)
load("UNL01_proc_data.RData")
# function to force scientific formatting of numbers (used for large BFs)
changeSciNot <- function(n) {
  output <- format(n, scientific = TRUE, digits = 2) #Transforms the number into scientific notation even if small
  output <- sub("e", "x10^", output) #Replace e with 10^
  output <- sub("\\+0?", "", output) #Remove + symbol and leading zeros on exponent, if > 1
  output <- sub("-0?", "-", output) #Leaves - symbol but removes leading zeros on exponent, if < 1
  output <- paste0(output,"^")
  # output = strsplit(output, "^", fixed = TRUE)
  # output = paste0(output[[1]][1],"^", output[[1]][2], "^")
  output
}

# function to extract and report BFs with error %s
report_BF_and_error <- function(BF_in, sci_not = TRUE, hyp = "alt"){
  
  if (hyp == "alt") {
    BF_notation = "BF~10~ = "
  } else if (hyp == "null") {
    BF_notation = "BF~01~ = "
  }
  
  if (sci_not == TRUE) {
    BF_value = changeSciNot(extractBF(BF_in)$bf) # change to sci notation
  } else {
    BF_value = round(extractBF(BF_in)$bf,2) # otherwise round
  }
  
  paste0(BF_notation, 
         BF_value, 
         " &plusmn; ", 
         round(100*extractBF(BF_in)$error,2), 
         "%")
}

```

```{r, include=FALSE}
#Create probable response variable, as we are using the uncertain condition
training <- training %>%
  mutate(prob_response = case_when((stage == "stage1" | stage == "stage2") & cue1 == "A" & cue_o_mouse.clicked_name == "o1_image" ~ 1,
                                   (stage == "stage1" | stage == "stage2") & cue1 == "A" & cue_o_mouse.clicked_name == "o2_image" ~ 0, 
                                   (stage == "stage1" | stage == "stage2") & cue1 == "B" & cue_o_mouse.clicked_name == "o1_image" ~ 0,
                                   (stage == "stage1" | stage == "stage2") & cue1 == "B" & cue_o_mouse.clicked_name == "o2_image" ~ 1,
                                   stage == "stage3" & cue2 == "X" & cue_o_mouse.clicked_name == "o1_image" ~ 1,
                                   stage == "stage3" & cue2 == "X" & cue_o_mouse.clicked_name == "o2_image" ~ 0, 
                                   stage == "stage3" & cue2 == "Y" & cue_o_mouse.clicked_name == "o1_image" ~ 0,
                                   stage == "stage3" & cue2 == "Y" & cue_o_mouse.clicked_name == "o2_image" ~ 1))  

stage1 <- filter(training, block <= 6)
stage1 <- subset(stage1, select = -c(s2_trial, s3_trial))

stage2 <- filter(training, (block >6) &(block <= 10))
stage2 <- subset(stage2, select = -c(s1_trial, s3_trial))

stage3 <- filter(training, block > 10)
stage3 <- subset(stage3, select = -c(s2_trial, s3_trial))
```

```{r, include=FALSE}
#some exclusion criteria
block6 <- filter(stage1, block == 6) %>%
  group_by(pNum, condition) %>%
  summarise (mean_response = mean(prob_response, na.rm = TRUE))

low_acc <- filter(block6, mean_response < 0.6) 
low_acc <- low_acc$pNum

training <- filter(training, !pNum %in% low_acc)
stage1 <- filter(stage1, !pNum %in% low_acc)
stage2 <- filter(stage2, !pNum %in% low_acc)
stage3 <- filter(stage3, !pNum %in% low_acc)

N <- unique(stage1$pNum)
pNum_cond <- subset(training, select = c("pNum", "condition"))
pNum_cond <- unique(pNum_cond)
count(pNum_cond, condition)
```

# All training phase

```{r, include=FALSE}
#Plot Training accuracy
MA_training <- training %>% 
  group_by(block, stage, condition) %>% 
  summarise(mean_prob_response = mean(prob_response, na.rm = TRUE),
            se_accuracy = sd(prob_response, na.rm = TRUE)/sqrt(length(prob_response)))
```

```{r, echo=FALSE}
ggplot(MA_training) +
  geom_point(mapping = aes(x = block, y = mean_prob_response, color = condition)) +
  geom_line(mapping = aes(x = block, y = mean_prob_response, color = condition)) +
  geom_errorbar(aes(x= block, y = mean_prob_response, ymin = mean_prob_response-se_accuracy, ymax = mean_prob_response+se_accuracy), color = "black", width=.1,position=position_dodge(0.05)) +
  facet_grid(cols = vars(stage), space = "free_x", scales = "free_x") + 
  scale_x_continuous(name = "Block", breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15)) +
  scale_y_continuous(name="Proportion of probable responses", limits=c(0.5, 1)) +
  labs(title = "Mean proportion of probable responses for the first stage of the training phase")
```

```{r, include=FALSE}
#ANOVA
train <- training %>%
  group_by (pNum, block, condition) %>%
  summarise(mean_response = mean(prob_response, na.rm = TRUE))
train$pNum <- factor(train$pNum)
train$block <- factor(train$block)
train$condition <- factor(train$condition)
ANOVA_train <- aov_car(formula = mean_response ~ condition + Error(pNum/block), data = train)
print(ANOVA_train)

bay_ANOVA_train <- anovaBF(formula = mean_response ~ condition + block + pNum,
        data = data.frame(train),
        whichRandom = "pNum")
print(bay_ANOVA_train)
bay_ANOVA_train_int <- bay_ANOVA_train[4]/bay_ANOVA_train[3]
print(bay_ANOVA_train_int)
```

```{r, include = FALSE}
# Marginal means
ANOVA_train_cond <-emmeans(ANOVA_train, ~condition)
print(ANOVA_train_cond)
ANOVA_train_block <-emmeans(ANOVA_train, ~block)
print(ANOVA_train_cond)
train_interaction <- emmeans(ANOVA_train, ~ block|condition)
print(train_interaction)
train_interaction2 <- emmeans(ANOVA_train, ~ condition|block)
print(train_interaction2)
```

```{r, include = FALSE}
#Simple main effect of group
SME_group <- train %>%
  group_by(block) %>%
  anova_test(mean_response ~ condition) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
print(SME_group)
```

```{r, include=FALSE}
#Simple comparisons for group
pw_ANOVA <- pairs(train_interaction2, adj = "bonferroni")
pw_ANOVA <- summary(pw_ANOVA)
print(pw_ANOVA)
```

A mixed model ANOVA, with the between subjects factor *group* and the within subjects factor *block* found both main effects significant (Group: `r apa(ANOVA_train, effect = "condition")`, `r report_BF_and_error(bay_ANOVA_train[2])`; Block:`r apa(ANOVA_train, effect = "block" )`, `r report_BF_and_error(bay_ANOVA_train[1])`) and the interaction between them (`r apa(ANOVA_train, effect = "condition:block")`, `r report_BF_and_error(bay_ANOVA_train_int[1])`), all of them with extreme evidence for the alternative hypothesis. Simple main effects showed an effect of the group in the blocks 2-7 (*F*(`r SME_group[3,3]` , `r SME_group[3,4]`) \> `r SME_group[3,5]`, *p* \< `r SME_group[3,9]`). In all the cases, the differences were between group U-U with the C-C and the C-U groups.

# Stage 1

```{r, include=FALSE}
#ANOVA
resp_s1 <- stage1 %>%
  group_by (pNum, block, condition) %>%
  summarise(mean_response = mean(prob_response, na.rm = TRUE))
resp_s1$block <- factor(resp_s1$block)
resp_s1$condition <- factor(resp_s1$condition)
resp_s1$pNum <- factor(resp_s1$pNum)
ANOVA_resp_s1 <- aov_car(formula = mean_response ~ condition + Error(pNum/block), data = resp_s1)
print(ANOVA_resp_s1)

bay_ANOVA_resp_s1 <- anovaBF(formula = mean_response ~ condition + block + pNum,
        data = data.frame(resp_s1),
        whichRandom = "pNum")
print(bay_ANOVA_resp_s1)
bay_ANOVA_resp_s1_int <- bay_ANOVA_resp_s1[4]/bay_ANOVA_resp_s1[3]
print(bay_ANOVA_resp_s1_int)
```

```{r, include = FALSE}
#Simple main effect of group
SME_ANOVAs2_group <- resp_s1 %>%
  group_by(block) %>%
  anova_test(mean_response ~ condition) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
print(SME_ANOVAs2_group)
```

```{r, include = FALSE}
# Marginal means
ANOVA_s1_cond <-emmeans(ANOVA_resp_s1, ~condition|block)
pairs(ANOVA_s1_cond, adj = "bonferroni")
```

In stage 1, all groups showed a similar increase in accuracy as blocks progressed. A mixed methods ANOVA confirmed a significant effect of the *block* (`r apa(ANOVA_resp_s1, effect = "block")`, `r report_BF_and_error(bay_ANOVA_resp_s1[1])`), the *condition* (`r apa(ANOVA_resp_s1, effect = "condition" )`, `r report_BF_and_error(bay_ANOVA_resp_s1[2])`), and of their interaction (`r apa(ANOVA_resp_s1, effect = "condition:block")`, `r report_BF_and_error(bay_ANOVA_resp_s1_int[1])`). simple main effect of condition showed that it was significant in all blocks except for the first one. Again, the differences were between group U-U with the other two groups.

# Stage 2

```{r, include=FALSE}
#ANOVA
resp_s2 <- stage2 %>%
  group_by (pNum, block, condition) %>%
  summarise(mean_response = mean(prob_response, na.rm = TRUE))
resp_s2$block <- factor(resp_s2$block)
resp_s2$condition <- factor(resp_s2$condition)
resp_s2$pNum <- factor(resp_s2$pNum)
ANOVA_resp_s2 <- aov_car(formula = mean_response ~ condition + Error(pNum/block), data = resp_s2)
print(ANOVA_resp_s2)

bay_ANOVA_resp_s2 <- anovaBF(formula = mean_response ~ condition + block + pNum,
        data = data.frame(resp_s2),
        whichRandom = "pNum")
print(bay_ANOVA_resp_s2)
bay_ANOVA_resp_s2_int <- bay_ANOVA_resp_s2[4]/bay_ANOVA_resp_s2[3]
print(bay_ANOVA_resp_s2_int)
```

```{r, include = FALSE}
#Simple main effect of block
SME_ANOVAs2_block <- resp_s2 %>%
  group_by(condition) %>%
  anova_test(mean_response ~ Error(pNum/block)) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
print(SME_ANOVAs2_block)

#Simple main effect of group
SME_ANOVAs2_group <- resp_s2 %>%
  group_by(block) %>%
  anova_test(mean_response ~ condition) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
print(SME_ANOVAs2_group)
```

```{r, include=FALSE}
#Simple comparisons for group
ANOVA_s2_cond <-emmeans(ANOVA_resp_s2, ~condition|block)
pw_ANOVA_s2 <- pairs(ANOVA_s2_cond, adj = "bonferroni")
pw_ANOVA_s2 <- summary(pw_ANOVA_s2)
print(pw_ANOVA_s2)
```

In stage 2, the C-C group continued with a high accuracy, and group U-U, kept increasing in accuracy. However, group C-U showed a decrease in the accuracy, to levels similar to the U-U group. A mixed model ANOVA did not find a significant effect of the *block* (`r apa(ANOVA_resp_s2, effect = "block" )`, `r report_BF_and_error(bay_ANOVA_resp_s2[1])`), but both the main effect of the *group* and the*group x block* interaction were significant (Group: `r apa(ANOVA_resp_s2, effect = "condition")`, `r report_BF_and_error(bay_ANOVA_resp_s2[2])`; Interaction: `r apa(ANOVA_resp_s2, effect = "condition:block")`, `r report_BF_and_error(bay_ANOVA_resp_s2_int[1])`). Simple main effects showed an effect of the condition in all blocks except for block 9. Simple main effects of condition showed that there were significant differences in block 7, but not in any more of them. Simple comparisons showed that, in block 7, group U-U was significantly different from group C-C (*t*(`r pw_ANOVA_s2[2, 5]`) = `r pw_ANOVA_s2[2, 6]`, *p* = `r pw_ANOVA_s2[2, 7]`) and group C-U (*t*(`r pw_ANOVA_s2[3, 5]`) = `r pw_ANOVA_s2[3, 6]`, *p* = `r pw_ANOVA_s2[3, 7]`).

# Stage 3

```{r, include=FALSE}
#ANOVA
resp_s3 <- stage3 %>%
  group_by (pNum, block, condition) %>%
  summarise(mean_response = mean(prob_response, na.rm = TRUE))
resp_s3$block <- factor(resp_s3$block)
resp_s3$condition <- factor(resp_s3$condition)
resp_s3$pNum <- factor(resp_s3$pNum)
ANOVA_s3 <- aov_car(formula = mean_response ~ condition + Error(pNum/block), data = resp_s3)
print(ANOVA_s3)

bay_ANOVA_s3 <- anovaBF(formula = mean_response ~ condition + block + pNum,
        data = data.frame(resp_s3),
        whichRandom = "pNum")
print(bay_ANOVA_s3)
bay_ANOVA_s3_int <- bay_ANOVA_s3[4]/bay_ANOVA_s3[3]
print(bay_ANOVA_s3_int)
```

```{r, include = FALSE}
# Marginal means
ANOVA_s3_block <-emmeans(ANOVA_s3, ~block)
pw_ANOVA_S3_block <- pairs(ANOVA_s3_block, adj = "bonferroni")
pw_ANOVA_S3_block<- summary(pw_ANOVA_S3_block)
print(pw_ANOVA_S3_block)
```

On the third stage, the ANOVA showed only a significant effect of the *block* (`r apa(ANOVA_s3, effect = "block" )`, `r report_BF_and_error(bay_ANOVA_s3[1])`), but not of the *condition* (`r apa(ANOVA_s3, effect = "condition" )`, `r report_BF_and_error(bay_ANOVA_s3[2])`), nor of their interaction (`r apa(ANOVA_s3, effect = "condition:block")`, `r report_BF_and_error(bay_ANOVA_s3_int[1])`).

## Stage 3 frecuent trials

```{r, include=FALSE}
#select the rare trials
stage3_frecuent <- filter(stage3, !((cue1 == "A") & (cue2 == "Y")) | ((cue1 == "B") & (cue2 == "X")))
```

```{r, include=FALSE}
M_s3_frec <- stage3_frecuent %>% 
  group_by(condition, block) %>%
  summarise(mean_acc = mean(correct_answer, na.rm = TRUE),
            se_accuracy = sd(correct_answer, na.rm = TRUE)/sqrt(length(correct_answer)))
```

```{r, echo=FALSE}
ggplot(M_s3_frec) +
  geom_point(mapping = aes(x = block, y = mean_acc, color = condition)) +
  geom_line(mapping = aes(x = block, y = mean_acc, color = condition)) +
  geom_errorbar(aes(x= block, y = mean_acc, ymin = mean_acc-se_accuracy, ymax = mean_acc+se_accuracy), color = "black", width=.1,position=position_dodge(0.05)) + 
  scale_x_continuous(name = "Block") +
  scale_y_continuous(name="Proportion of probable responses") +
  labs(title = "Mean proportion of probable responses for the frecuent pairs in Stage 3")
```

```{r, include=FALSE}
#ANOVA
resp_s3_frec <- stage3_frecuent %>%
  group_by (pNum, block, condition) %>%
  summarise(mean_response = mean(prob_response, na.rm = TRUE))
resp_s3_frec$block <- factor(resp_s3_frec$block)
resp_s3_frec$condition <- factor(resp_s3_frec$condition)
resp_s3_frec$pNum <- factor(resp_s3_frec$pNum)
ANOVA_s3_frec <- aov_car(formula = mean_response ~ condition + Error(pNum/block), data = resp_s3_frec)
print(ANOVA_s3_frec)

bay_ANOVA_s3_frec <- anovaBF(formula = mean_response ~ condition + block + pNum,
        data = data.frame(resp_s3_frec),
        whichRandom = "pNum")
print(bay_ANOVA_s3_frec)
bay_ANOVA_s3_frec_int <- bay_ANOVA_s3_frec[4]/bay_ANOVA_s3_frec[3]
print(bay_ANOVA_s3_frec_int)
```

On the frequent trials (AY-O2, BX - O1) in phase three, the ANOVA showed only a significant effect of the *block* (`r apa(ANOVA_s3_frec, effect = "block" )`, `r report_BF_and_error(bay_ANOVA_s3_frec[1])`), but not of the *condition* (`r apa(ANOVA_s3_frec, effect = "condition" )`, `r report_BF_and_error(bay_ANOVA_s3_frec[2])`), but not of their interaction (`r apa(ANOVA_s3_frec, effect = "condition:block")`, `r report_BF_and_error(bay_ANOVA_s3_frec_int[1])`).

## Stage 3 rare trials

```{r, include=FALSE}
#select the rare trials
stage3_rare <- filter(stage3, ((cue1 == "A") & (cue2 == "Y")) | ((cue1 == "B") & (cue2 == "X")))
```

```{r, include=FALSE}
M_s3_rare <- stage3_rare %>% 
  group_by(condition, block) %>%
  summarise(mean_acc = mean(correct_answer, na.rm = TRUE),
            se_accuracy = sd(correct_answer, na.rm = TRUE)/sqrt(length(correct_answer)))
```

```{r, echo=FALSE}
ggplot(M_s3_rare) +
  geom_point(mapping = aes(x = block, y = mean_acc, color = condition)) +
  geom_line(mapping = aes(x = block, y = mean_acc, color = condition)) +
  geom_errorbar(aes(x= block, y = mean_acc, ymin = mean_acc-se_accuracy, ymax = mean_acc+se_accuracy), color = "black", width=.1,position=position_dodge(0.05)) + 
  scale_x_continuous(name = "Block") +
  scale_y_continuous(name="Proportion of probable responses") +
  labs(title = "Mean proportion of probable responses for the infrecuent pairs in Stage 3")
```

```{r, include=FALSE}
#ANOVA
resp_s3_rare <- stage3_rare %>%
  group_by (pNum, block, condition) %>%
  summarise(mean_response = mean(prob_response, na.rm = TRUE))
resp_s3_rare$block <- factor(resp_s3_rare$block)
resp_s3_rare$condition <- factor(resp_s3_rare$condition)
resp_s3_rare$pNum <- factor(resp_s3_rare$pNum)
ANOVA_s3_rare <- aov_car(formula = mean_response ~ condition + Error(pNum/block), data = resp_s3_rare)
print(ANOVA_s3_rare)

bay_ANOVA_s3_rare <- anovaBF(formula = mean_response ~ condition + block + pNum,
        data = data.frame(resp_s3_rare),
        whichRandom = "pNum")
print(bay_ANOVA_s3_rare)
bay_ANOVA_s3_rare_int <- bay_ANOVA_s3_rare[4]/bay_ANOVA_s3_rare[3]
print(bay_ANOVA_s3_rare_int)
```

```{r, include = FALSE}
#Simple main effect of block
SME_s3_rare_block <- resp_s3_rare %>%
  group_by(condition) %>%
  anova_test(mean_response ~ Error(pNum/block)) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
print(SME_s3_rare_block)

#Simple main effect of group
SME_s3_rare_group <- resp_s3_rare %>%
  group_by(block) %>%
  anova_test(mean_response ~ condition) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
print(SME_s3_rare_group)
```

```{r, include=FALSE}
#Simple comparisons for group
ANOVA_s3_rare_cond <-emmeans(ANOVA_s3_rare, ~condition|block)
pw_ANOVA_s3_rare <- pairs(ANOVA_s3_rare_cond, adj = "bonferroni")
pw_ANOVA_s3_rare <- summary(pw_ANOVA_s3_rare)
print(pw_ANOVA_s3_rare)
```

On the infrequent trials (AY-O2, BX - O1) in phase three, the ANOVA showed only a significant effect of the *block* (`r apa(ANOVA_s3_rare, effect = "block" )`, `r report_BF_and_error(bay_ANOVA_s3_rare[1])`), but not of the *condition* (`r apa(ANOVA_s3_rare, effect = "condition" )`, `r report_BF_and_error(bay_ANOVA_s3_rare[2])`). The interaction was significant (`r apa(ANOVA_s3_rare, effect = "condition:block")`, `r report_BF_and_error(bay_ANOVA_s3_rare_int[1])`). However, simple main effects showed no effect of the condition in any of the blocks.
